---
title: "Sensitivity Models: Sarilumab Meta-Regressions"
author: "Arthur M. Albuquerque"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      fig.align = 'center')
```

```{r message=FALSE, warning=FALSE}
# Ensures the package "pacman" is installed
if (!require("pacman")) install.packages("pacman")

pacman::p_load(here,
               dplyr,
               ggplot2,
               bayesmeta,
               brms,
               metafor)

remotes::install_github("stan-dev/cmdstanr")

# Ensure that packages are installed according to versions in renv's lockfile
renv::restore()

# Load data
d_logOR = readRDS(
  here::here("output", "data", "effect_sizes.Rds")
)
```


## Tocilizumab vs. Sarilumab

Here are the sensitivity analyses priors. Here, we used the results from REMAP-CAP to define the SDs for both "Optimistic for Sarilumab" and "Optimistic for Tocilizumab" priors:

$$
\begin{align*}
y_i & \sim Normal(\theta_i, \sigma_i^2) \tag{Likelihood}\\
\theta_i & \sim Normal(\mu, \tau^2)\\
\mu &= \beta_0 + \beta_1 x_i\\
\\
\beta_0 & \sim \operatorname{Normal}(0, 0.75^2) \tag{Priors}\\
\tau & \sim \operatorname{Log-Normal}(-1.975, 0.67^2)\\
\beta_{1[Sarilumab]} & \sim \operatorname{Normal}(-0.049, 0.118^2)\\
\beta_{1[Tocilizumab]} & \sim \operatorname{Normal}(0.049, 0.118^2)\\
\end{align*}
$$

```{r}
## Means
# Source: Figure S3, Supplementary Appendix page 46 
# https://www.medrxiv.org/content/10.1101/2021.06.18.21259133v2.supplementary-material
# version posted June 22, 2021
reciprocal_remap_cap = 1/1.05
mean_log_sari = log(reciprocal_remap_cap)
## SDs
# Formula to extract standard error (SE)
# Transform the results into the log scale and calculate the SE
# https://training.cochrane.org/handbook/current/chapter-06#section-6-3-2
width = log(1.35) - log(0.85)
sari_sd_optimistic_sari = width/(2*1.96)
sari_sd_optimistic_toci = sari_sd_optimistic_sari
```

```{r}
# General model characteristics 
## Formula
mf = 
  formula(yi | se(sei) ~ 0 + Intercept + treatment + (1 | study))
## Priors for beta_0 and tau
informative = bayesmeta::TurnerEtAlPrior("all-cause mortality",
                                         "pharmacological",
                                         "placebo / control")
logmean = informative$parameters["tau", "mu"]
logsd = informative$parameters["tau", "sigma"]
# tau:
# logmean
# -1.975
# logsd
# 0.67
priors = 
  brms::prior(normal(0, 0.75), class = "b", coef = "Intercept") +
  brms::prior(lognormal(-1.975, 0.67), class = "sd", group = "study")
```

Fit models

```{r}
## "Optimistic for Sarilumab" model
# mean_log_sari
# -0.04879016
# sari_sd_optimistic_sari
# 0.1180162
m_sari_optimistic_sari_remapcap =
  brms::brm(
    
  data = d_logOR |> dplyr::filter(treatment != "baricitinib"),
  family = gaussian,
  formula = mf,
  
  prior = priors +
    brms::prior(normal(-0.04879016, 0.1180162), class = "b", coef = "treatmentsarilumab"),
  sample_prior = TRUE,
  
  control = list(adapt_delta = .96),
  backend = "cmdstanr", # faster
  warmup = 2000,
  iter = 4000,
  seed = 123,
  cores = parallel::detectCores(),
  chains = 4,
  
  file = here::here("output", "fits", "sensitivity_analyses",
                    "mr_sari_optimistic_sari_remapcap.Rds"),
  file_refit = "on_change"
)
## "Optimistic for Tocilizumab" model
# -mean_log_sari
# 0.04879016
# sari_sd_optimistic_toci
# 0.1180162
m_sari_optimistic_toci_remapcap =
  brms::brm(
    
  data = d_logOR |> dplyr::filter(treatment != "baricitinib"),
  family = gaussian,
  formula = mf,
  
  prior = priors +
    brms::prior(normal(0.04879016, 0.1180162), class = "b", coef = "treatmentsarilumab"),
  sample_prior = TRUE,
  
  control = list(adapt_delta = .96),
  backend = "cmdstanr", # faster
  warmup = 2000,
  iter = 4000,
  seed = 123,
  cores = parallel::detectCores(),
  chains = 4,
  
  file = here::here("output", "fits", "sensitivity_analyses",
                    "mr_sari_optimistic_toci_remapcap.Rds"),
  file_refit = "on_change"
)
```

## Tocilizumab vs. Baricitinib

Here are the sensitivity analyses priors. Here, we used the results from 
Karampitsakos et al. (DOI 10.1101/2022.06.13.22276211) to define the SDs for
both "Optimistic for Baricitinib" and "Optimistic for Tocilizumab" priors:

$$
\begin{align*}
y_i & \sim Normal(\theta_i, \sigma_i^2) \tag{Likelihood}\\
\theta_i & \sim Normal(\mu, \tau^2)\\
\mu &= \beta_0 + \beta_1 x_i\\
\\
\beta_0 & \sim \operatorname{Normal}(0, 0.75^2) \tag{Priors}\\
\tau & \sim \operatorname{Log-Normal}(-1.975, 0.67^2)\\
\beta_{1[Baricitinib]} & \sim \operatorname{Normal}(-0.335, 0.264^2)\\
\beta_{1[Tocilizumab]} & \sim \operatorname{Normal}(0.335, 0.264^2)\\
\end{align*}
$$

```{r}
# Data provided by Karampitsakos et al. through email 

# Calculate crude log odds ratio

bari_toci_data = 
  metafor::escalc(
  measure = "OR", # log odds ratio
  
  # Baricitinib
  ai = 40,
  n1i = 125,
  
  # Tocilizumab
  ci = 50,
  n2i = 126
) %>%
  as_tibble() %>% 
  # Calculate standard error
  dplyr::mutate(sei = sqrt(vi))

## Mean
mean_log_bari = bari_toci_data$yi[1]

## SDs
bari_sd_optimistic_bari = bari_toci_data$sei[1]
bari_sd_optimistic_toci = bari_sd_optimistic_bari
```

```{r}
# General model characteristics 
## Formula
mf = 
  formula(yi | se(sei) ~ 0 + Intercept + treatment + (1 | study))
## Priors for beta_0 and tau
informative = bayesmeta::TurnerEtAlPrior("all-cause mortality",
                                         "pharmacological",
                                         "placebo / control")
logmean = informative$parameters["tau", "mu"]
logsd = informative$parameters["tau", "sigma"]
# tau:
# logmean
# -1.975
# logsd
# 0.67
priors = 
  brms::prior(normal(0, 0.75), class = "b", coef = "Intercept") +
  brms::prior(lognormal(-1.975, 0.67), class = "sd", group = "study")
```

Fit models

```{r}
## "Optimistic for Baricitinib" model
# mean_log_bari
# -0.3350615
# bari_sd_optimistic_bari
# 0.2644288

m_bari_optimistic_bari_direct =
  brms::brm(
  
  # only include RECOVERY Bari without patients treated with Tocilizumab  
  data = d_logOR |> dplyr::filter(treatment != "sarilumab", study != "RECOVERY Bari"),
  family = gaussian,
  formula = mf,
  
  prior = priors +
    brms::prior(normal(-0.3350615, 0.2644288), class = "b", coef = "treatmentbaricitinib"),
  sample_prior = TRUE,
  
  control = list(adapt_delta = .96),
  backend = "cmdstanr", # faster
  warmup = 2000,
  iter = 4000,
  seed = 123,
  cores = parallel::detectCores(),
  chains = 4,
  
  file = here::here("output", "fits", "sensitivity_analyses",
                    "mr_bari_optimistic_bari_direct.Rds"),
  file_refit = "on_change"
)
## "Optimistic for Tocilizumab" model
# -mean_log_bari
# 0.3350615
# bari_sd_optimistic_bari
# 0.2644288

m_bari_optimistic_toci_direct =
  brms::brm(
  
  # only include RECOVERY Bari without patients treated with Tocilizumab    
  data = d_logOR |> dplyr::filter(treatment != "sarilumab", study != "RECOVERY Bari"),
  family = gaussian,
  formula = mf,
  
  prior = priors +
    brms::prior(normal(0.3350615, 0.2644288), class = "b", coef = "treatmentbaricitinib"),
  sample_prior = TRUE,
  
  control = list(adapt_delta = .96),
  backend = "cmdstanr", # faster
  warmup = 2000,
  iter = 4000,
  seed = 123,
  cores = parallel::detectCores(),
  chains = 4,
  
  file = here::here("output", "fits", "sensitivity_analyses",
                    "mr_bari_optimistic_toci_direct.Rds"),
  file_refit = "on_change"
)
```

# Model diagnostics

```{r}
# Load function
source(here::here("functions", "diag_plot.R"))
```

## Tocilizumab vs. Sarilumab

"Optimistic for Sarilumab" model

```{r}
diag_plot(model = m_sari_optimistic_sari_remapcap,
          pars_list = c("b_Intercept", "b_treatmentsarilumab", "sd_study__Intercept"),
          ncol_trace = 4)
```

```{r message=FALSE, warning=FALSE}
brms::pp_check(m_sari_optimistic_sari_remapcap,
               type = "dens_overlay_grouped",
               group = "treatment",
               ndraws = 50)

brms::pp_check(m_sari_optimistic_sari_remapcap,
               type = "ecdf_overlay_grouped",
               group = "treatment",
               ndraws = 50)
```

"Optimistic for Tocilizumab" model

```{r}
diag_plot(model = m_sari_optimistic_toci_remapcap,
          pars_list = c("b_Intercept", "b_treatmentsarilumab", "sd_study__Intercept"),
          ncol_trace = 4)
```

```{r message=FALSE, warning=FALSE}
brms::pp_check(m_sari_optimistic_toci_remapcap,
               type = "dens_overlay_grouped",
               group = "treatment",
               ndraws = 50) 

brms::pp_check(m_sari_optimistic_toci_remapcap,
               type = "ecdf_overlay_grouped",
               group = "treatment",
               ndraws = 50)
```

## Tocilizumab vs. Baricitinib

"Optimistic for Baricitinib" model

```{r}
diag_plot(model = m_bari_optimistic_bari_direct,
          pars_list = c("b_Intercept", "b_treatmentbaricitinib", "sd_study__Intercept"),
          ncol_trace = 4)
```

```{r message=FALSE, warning=FALSE}
brms::pp_check(m_bari_optimistic_bari_direct,
               type = "dens_overlay_grouped",
               group = "treatment",
               ndraws = 50)

brms::pp_check(m_bari_optimistic_bari_direct,
               type = "ecdf_overlay_grouped",
               group = "treatment",
               ndraws = 50)
```

"Optimistic for Tocilizumab" model

```{r}
diag_plot(model = m_bari_optimistic_toci_direct,
          pars_list = c("b_Intercept", "b_treatmentbaricitinib", "sd_study__Intercept"),
          ncol_trace = 4)
```

```{r message=FALSE, warning=FALSE}
brms::pp_check(m_bari_optimistic_toci_direct,
               type = "dens_overlay_grouped",
               group = "treatment",
               ndraws = 50) 

brms::pp_check(m_bari_optimistic_toci_direct,
               type = "ecdf_overlay_grouped",
               group = "treatment",
               ndraws = 50)
```

